---
params:
    School: "Cramp, William Elementary"
    StartDate: "08/01/2019"
    EndDate: "10/01/2019"
title: "STEP Metrics: August 2019 - September 2019"
author: "Jessica Streeter, Data Scientist | Data Informatics, CBH"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
  pdf_document: 
    latex_engine: xelatex
classoption: landscape
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}


---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
knitr::opts_chunk$set(fig.width = 11, fig.height = 7.5, fig.align = "center")

options(tinytex.verbose = TRUE) #long form errors issued for latex
options(kableExtra.latex.load_packages = FALSE) #need this to get kableextra and tinytex to play nice
options(knitr.table.format = "latex")


#School Parameter####
#SchoolNum <- 20
#SchoolName <- unique(STEPfinal$SchoolName[as.numeric(STEPfinal$SchoolName) == SchoolNum]) - this is defined at end bc code runs in order


#Libraries####
library(tidyverse)
library(data.table)
library(zoo)
library(kableExtra)
library(lubridate)
library(magrittr)


#reading in data####
#streamlined version of the code in STEP data cleaning


#reading data####
STEPfinal <- readRDS(paste0("STEPfinal_", format(zoo::as.yearmon(Sys.Date()), format = "%b-%Y"))) 

STEPfinal <- STEPfinal %>% filter(EncounterDate <= as.Date(params$EndDate, format = "%m/%d/%Y"),
                                  EncounterDate >= as.Date(params$StartDate, format = "%m/%d/%Y"))

```
# School: `r params$School`
## Address: `r unique(STEPfinal$Address1[STEPfinal$SchoolName == params$School])`

#### Summary Statistics
All data from `r format(as.Date(params$StartDate, format = "%m/%d/%Y"), "%B %Y")` through `r format(as.Date(params$EndDate, format = "%m/%d/%Y"), "%B %Y")`

* Total Number of School-Specific Activities: `r nrow(STEPfinal[STEPfinal$SchoolName == params$School & is.na(STEPfinal$TaskType) == F & STEPfinal$TaskCat == "School" & STEPfinal$EncounterDate <= as.Date(params$EndDate, "%m/%d/%Y"),])` 
* Total Number of Child-Specific Activities: `r nrow(STEPfinal[STEPfinal$SchoolName == params$School & is.na(STEPfinal$TaskType) == F & STEPfinal$TaskCat == "Child" & STEPfinal$EncounterDate <= as.Date(params$EndDate, "%m/%d/%Y"),])` 
* Number of Unique Children Served: `r length(unique(STEPfinal$ChildID[STEPfinal$SchoolName == params$School & is.na(STEPfinal$TaskType) == F & STEPfinal$TaskCat == "Child" & STEPfinal$EncounterDate <= as.Date(params$EndDate, "%m/%d/%Y")]))`

#### Data Notes

* All data reported by STEP programs through form.com.
* The number of distinct children is an approximate, as some data points include more than one child.
* Elapsed times that were less than zero were removed, as they were likely data entry errors.
* There were instances of tasks being entered with no times or incomplete time information.  These data points were removed. 
* Data with no encounter date or school name were removed.
* Collateral contacts with family, teachers, and other supports were grouped into "Collateral Contacts" for figures only.
* Collateral family therapy, family therapy with child, group therapy, and individual therapy were grouped into "Therapy" for figures only.

\newpage
```{r FIG1 School Metrics}
#School Specific Metric Figures####
STEPfinal %>% 
  filter(TaskCat == "School") %>% 
  mutate(YMencounter = as.yearmon(EncounterDate),
         TaskType_Cat = as.factor(ifelse(TaskType %in% c("Collateral Contact with Family", "Collateral Contact with Other Supports",
                                                         "Collateral Contact with Teacher"), 'Collateral Contacts',
                                         ifelse(TaskType %in% c("Collateral Family Therapy (Without Child)", "Family Therapy w/Child", 
                                                                "Group Therapy", "Individual Therapy w/Child"), "Therapy", as.character(TaskType))))) %>% 
  group_by(SchoolName, YMencounter, TaskCat, TaskType_Cat) %>% 
  summarize(TotalTime = round(sum(TimeElapsed)/3600,2), #Note: 4 cases where timeelapsed is NA
            TotalEntries = n(),
            NDistinctKids = n_distinct(ChildID)) %>% 
  filter(SchoolName == params$School) %>% 
  ggplot(aes(x = YMencounter, y = TotalTime, group = TaskType_Cat, fill = TaskType_Cat)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_label(aes(label = paste0(TotalTime, "H; ", "N = ", TotalEntries)), 
             position = position_stack(vjust = .5), 
             size = 3, color = "black", alpha = .3, fill = "white", 
             show.legend = NA, label.size = 0, label.padding = unit(.1, "lines")) +
  scale_x_yearmon(breaks = seq(from = as.yearmon(as.Date(params$StartDate, format = "%m/%d/%Y")), 
                               to = as.yearmon(as.Date(params$EndDate, format = "%m/%d/%Y")),
                               by = 1/12)) +
  #scale_y_time(breaks = seq(0,48*60*60, 8*60*60)) + #bc data is stored in seconds
  scale_fill_viridis_d(option = "D") +
  #theme(axis.text.x = element_text(angle= 45, hjust = 1)) +
  labs(title = "Time Spent on School-Specific Tasks",
       subtitle = paste0("School: ", params$School),
       x = "Month of Service",
       y = "Total Time (Hours)",
       fill = "Task Category",
       caption = "N: Total Number of Encounters") +
  theme(legend.position = "bottom")
```
\newpage
```{r FIG2 Child Metrics}
#Child Specific Metric Figures####
STEPfinal %>% 
  filter(TaskCat == "Child") %>% 
  mutate(YMencounter = as.yearmon(EncounterDate),
         TaskType_Cat = as.factor(ifelse(TaskType %in% c("Collateral Contact with Family", "Collateral Contact with Other Supports",
                                                  "Collateral Contact with Teacher"), 'Collateral Contacts',
                                  ifelse(TaskType %in% c("Collateral Family Therapy (Without Child)", "Family Therapy w/Child", 
                                                         "Group Therapy", "Individual Therapy w/Child"), "Therapy", as.character(TaskType))))) %>% 
  group_by(SchoolName, YMencounter, TaskCat, TaskType_Cat) %>% 
  summarize(TotalTime = round(sum(TimeElapsed)/3600,1), #Note: 4 cases where timeelapsed is NA
            TotalEntries = n(),
            NDistinctKids = n_distinct(ChildID)) %>% 
  filter(SchoolName == params$School) %>% 
  ggplot(aes(x = YMencounter, y = TotalTime, group = TaskType_Cat, fill = TaskType_Cat)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_label(aes(label = paste0(TotalTime,"H; N = ", NDistinctKids)), 
            position = position_stack(vjust = .5), 
            size = 2, color = "black", alpha = .3, fill = "white", 
            show.legend = NA, label.size = 0, label.padding = unit(.1, "lines")) +
  scale_x_yearmon(breaks = seq(from = as.yearmon(as.Date(params$StartDate, format = "%m/%d/%Y")), 
                               to = as.yearmon(as.Date(params$EndDate, format = "%m/%d/%Y")),
                               by = 1/12)) +
  #scale_y_time(breaks = seq(0,48*60*60, 8*60*60)) + #bc data is stored in seconds
  scale_fill_viridis_d(option = "E") +
  #theme(axis.text.x = element_text(angle= 45, hjust = 1)) +
  labs(title = "Time Spent on Child-Specific Tasks",
       subtitle = paste0("School: ", params$School),
       x = "Month of Service",
       y = "Total Time (Hours)",
       fill = "Task Category",
       caption = "'Therapy' and 'Collateral Contacts' were aggregated from more detailed data\n
       N: Number of Unique Kids") +
  theme(legend.position = "bottom")

```

\newpage
```{r Table1 SUmmary Data}
#summarytable for overall monthly hours.

STEPfinal %>% 
  mutate(YMencounter = as.yearmon(EncounterDate),
         TimeElapsed = as.ITime(TimeElapsed)) %>% 
  group_by(SchoolName, YMencounter, TaskCat) %>% 
  summarize(TotalTime = round(sum(TimeElapsed)/3600,2), #Note: 4 cases where timeelapsed is NA
            TotalEntries = n(),
            NDistinctKids = n_distinct(ChildID)) %>% 
  mutate(NDistinctKids = ifelse(TaskCat == "Child", NDistinctKids, NA)) %>% 
  filter(SchoolName == params$School) %>% #, YMencounter == MonthParam) %>% 
  arrange(YMencounter) %>% 
  ungroup() %>% 
  select(YMencounter, TaskCat, TotalTime, TotalEntries, NDistinctKids) %>%  #for specifying row groupings
  mutate(NDistinctKids = cell_spec(NDistinctKids, "latex", color = ifelse(is.na(NDistinctKids) == T, "white", "black"), align = "r")) %>% 
  rename("Encounter Month" = YMencounter,
          "Total Time(H)" = TotalTime,
          "Task Group" = TaskCat,
         "Total Entries" = TotalEntries,
         "N Unique Children" = NDistinctKids) %>% 
  kable(booktabs = T, escape = F, linesep = "",
        caption = paste0("Monthly Summary Data for ", params$School)) %>% 
  kable_styling(latex_options = c("condensed", "striped"), full_width = F) %>%
  column_spec(1, bold = T, italic = T) %>% 
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle")

```

```{r Table2 Detailed Data}
#SummaryTable 
STEPfinal %>% 
  mutate(YMencounter = as.yearmon(EncounterDate)) %>% 
  group_by(SchoolName, YMencounter, TaskCat, TaskType) %>% 
  summarize(TotalTime = round(sum(TimeElapsed)/3600,2), #Note: 4 cases where timeelapsed is NA
            TotalEntries = n(),
            NDistinctKids = n_distinct(ChildID)) %>% 
  mutate(NDistinctKids = ifelse(TaskCat == "Child", NDistinctKids, NA)) %>% 
  filter(SchoolName == params$School) %>% #, YMencounter == MonthParam) %>%
  spread(key = TaskCat, value = TotalTime) %>% 
  arrange(YMencounter, School, TaskType) %>% 
  ungroup() %>% 
  select(YMencounter, TaskType, School,Child,TotalEntries, NDistinctKids) %>% #for specifying row groupings
  mutate(School = cell_spec(School, "latex", color = ifelse(is.na(School) == T, "white", "black"), align = "r"),
         Child = cell_spec(Child, "latex", color = ifelse(is.na(Child) == T, "white", "black"), align = "r"),
         NDistinctKids = cell_spec(NDistinctKids, "latex", color = ifelse(is.na(NDistinctKids) == T, "white", "black"), align = "r")) %>% 
  #select(TaskType_Cat, School,Child,TotalEntries, NDistinctKids) %>% 
  rename("        " = YMencounter,
         Task = TaskType,
         "School: Total Time(H)" = School,
         "Child: Total Time(H)" = Child,
         "Total Entries" = TotalEntries,
         "N Unique Children" = NDistinctKids) %>%  
  kable(longtable = T, booktabs = T, escape = F, linesep = "", 
        caption = paste0("Detailed Data for ", params$School)) %>% 
  kable_styling(latex_options = c("scale_down", "striped", "repeat_header")) %>%
  column_spec(1, bold = TRUE, italic = TRUE) %>% 
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle")
  
```